---
title: "ARC Subaward Analysis"
author: "Nicholas Chen"
date: 2023-10-04
output: 
  bookdown::html_document2:
    fig_caption: yes
    number_sections: yes
    toc: yes
    reader-mode: true
    embed-resources: true
    code_folding: hide
number-sections: true
code-line-numbers: true
code-overflow: wrap
code-link: true
---

## Analysis of USAID Sub-award Spending

### Creating Dataframe

The following analysis walks through how to collate and analyze USASpending 
contracts data subawards. All functions used documented in roxygen.

Begin by loading in the data. Sample data used in this analysis is a complete 
data download from USASpending filtering for all contracts where 
"Awarding Agency" = USAID as of 10/1/2023.

```{r}
library(tidyverse)
prim_df <- read.csv("./data/Contracts_PrimeAwardSummaries_2023-10-01_H04M40S14_1.csv")
sub_df <- read.csv("./data/Contracts_Subawards_2023-10-01_H04M45S53_1.csv")
```


Clean and join the data so primary awards are matched with subaward data.
```{r}
source("./.R/subaward_df.R")
subaward_df(prim_df, sub_df, "JoinedData") ## df assigned to JoinedData
JoinedData %>%
  slice(1:10) ## view first ten rows
```
### Analyzing function

The function subaward_summary returns a dataframe of subaward metrics grouped 
by parent receiver.

```{r}
## list of countrycodes and awarding_office_codes
prim_df %>%
  group_by(awarding_office_code, awarding_office_name) %>%
  summarize() -> awarding_office_codes
prim_df %>%
  group_by(primary_place_of_performance_country_code, 
           primary_place_of_performance_country_name) %>%
  summarize() -> country_codes
View(awarding_office_codes)
View(country_codes)
```


The following returns a summarydf trimmed since the august 2010 directive.  Use
mmyydd for date filter.  The function uses the following parameters:

joined_df: joined_df from subaward_df
start_date: yymmdd.  All contracts before are excluded
contract_threshold: numeric.  All contracts with obligated amounts below excluded
country_code character: Filters for primary_place_of_performance_country_code
award_office_code: character. Filters for awarding_office_code
by_year: logical. If true, divides final summary table by year
df_name character: Name which you want df saved to

```{r}
source("./.R/subaward_summary.R")
subaward_summary(JoinedData, start_date = "2010-08-01", df_name = "after_mandate")
after_mandate %>%
  mutate(total_ob = scales::dollar_format()(total_ob)) %>%
  mutate(subawardedprojects_ob = scales::dollar_format()(subawardedprojects_ob)) %>%
  View()
```

After mandate excluding contracts below 25k

```{r}
subaward_summary(JoinedData, start_date = "2010-08-01", contract_threshold = 25000, df_name = "exclude_low_ctrcts")
exclude_low_ctrcts %>%
  mutate(total_ob = scales::dollar_format()(total_ob)) %>%
  mutate(subawardedprojects_ob = scales::dollar_format()(subawardedprojects_ob)) %>%
  View()
```


after mandate excluding low contracts subsetted by year

```{r}
subaward_summary(JoinedData, start_date = "2010-08-01", 
                 contract_threshold = 25000, 
                 df_name = "by_year",
                 by_year = T)
by_year %>%
  mutate(total_ob = scales::dollar_format()(total_ob)) %>%
  mutate(subawardedprojects_ob = scales::dollar_format()(subawardedprojects_ob)) %>%
  mutate(aggregate_obs = scales::dollar_format()(aggregate_obs)) %>%
  View()
```

Columbia specific

```{r}
subaward_summary(JoinedData, start_date = "2010-08-01", country_code = "COL", df_name = "col_data")
col_data %>%
  mutate(total_ob = scales::dollar_format()(total_ob)) %>%
  mutate(subawardedprojects_ob = scales::dollar_format()(subawardedprojects_ob)) %>%
  View()
```

